---
title: "Medical data analysis"
author: "James Porter as Burrbank"
date: "January 11, 2017"
output: html_document
---

First read the data into R

```{r}
library(feather)

names <- read_feather('data/drugnames.feather')
spend2011 <- read_feather('data/spending-2011.feather')
spend2012 <- read_feather('data/spending-2012.feather')
spend2013 <- read_feather('data/spending-2013.feather')
spend2014 <- read_feather('data/spending-2014.feather')
spend2015 <- read_feather('data/spending-2015.feather')

```

Questions for the data. 

I think my very first step is to sort the data on the generic drug name and year to create a larger data frame.

```{r}

# first add the year into each dataframe
spend2011$year <- 2011
spend2012$year <- 2012
spend2013$year <- 2013
spend2014$year <- 2014
spend2015$year <- 2015

spend <- rbind(spend2011, spend2012, spend2013, spend2014, spend2015)

spend <- test[order(test$drugname_generic, test$year),]

```

I want to do a bit of exploration to see what the general trend of drug spending was between 2011 and 2015, then maybe look at some sort of profile of more costly medications. Basically I just want to get a good idea of what the data looks like at this point maybe some more thoughts about what this data set could have in store will come.

I want to see what kind of drugs seem to cost the most/patient and which are costing the most in total.


Then I will do some graphing, I need to find some way too look at the data in some meaningful way.

