---
title: "atc_medicare_merge"
author: "darya akimova"
date: "February 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

The goal of this work is to merge the drug names from the Medicare Part D spending data with the ATC classifications system, based on drug names. Anticipating having to clean both sets of names to get them to merge.

Packages:
```{r packages}
library(tidyverse)
library(cowplot)
library(data.world)
```

Data:
```{r data}
drug.spend.ds <- "https://data.world/data4democracy/drug-spending"
# drug names from Medicare Part D data:
d.drugs <- data.world::query(
  data.world::qry_sql("SELECT DISTINCT brand_name, generic_name FROM `spending_part_d_2011to2015_tidy`"),
  dataset = drug.spend.ds
  ) %>% 
  mutate_all(str_trim)  # just in case
glimpse(d.drugs)
# atc classification system from drug_uses.csv
atc <- data.world::query(
  data.world::qry_sql("SELECT * FROM drug_uses"),
  dataset = drug.spend.ds
) 
glimpse(atc)
```

The ATC datasets needs some cleaning up:

```{r atc_cleanup}
atc <- atc %>% 
  select(-column_a) %>% # not useful
  select(starts_with("drug"), substance, name, everything()) %>% 
  mutate_all(str_to_lower) %>% 
  mutate_all(str_trim) %>% 
  # rename some variables for clarity
  rename(
    first_level = anatomical,
    second_level = therapeutic,
    third_level = pharmacologic,
    fourth_level = chemical
    )
glimpse(atc)
# it looks like drugname_generic and substance might be identical columns, is this true?
all.equal(atc$drugname_generic, atc$substance)  # yes
# remove the substance column, rename the `name` column appropriately to what it really is - the chemical formulation
atc <- atc %>%
  select(-substance) %>% 
  rename(
    chem_formula = name
  )
glimpse(atc)
# much better
glimpse(d.drugs)
```

Time for some test merges - the low hanging fruit is:
atc - `drugname_brand` to d.drugs - `brand_name`

```{r}
brand.semi <- d.drugs %>% 
  semi_join(atc, by = c("brand_name" = "drugname_brand"))
dim(brand.semi)
```

1241 drug matches our of 4498! Not bad at all.

Will take what is left over and keep working on it:
```{r}
brand.anti <- d.drugs %>% 
  anti_join(atc, by = c("brand_name" = "drugname_brand")) %>% 
  arrange(brand_name)
glimpse(brand.anti)
```

Future problems ahead include:
- instruments vs drugs (needles, the like)
- multiple drug ingredients in the generic_name column 

On and further
```{r}
generic.semi <- brand.anti %>% 
  semi_join(atc, by = c("generic_name" = "drugname_generic"))
dim(generic.semi)
# whump whump
chem.semi <- brand.anti %>% 
  semi_join(atc, by = c("generic_name" = "chem_formula"))
dim(chem.semi)
chem.anti <- brand.anti %>% 
  anti_join(atc, by = c("generic_name" = "chem_formula"))
dim(chem.anti)
```

That is 84 more than before, I suppose. A little less than what I was hoping for.

Break 1:

```{r}
matched.so.far <- brand.semi %>% 
  bind_rows(chem.semi)
dim(matched.so.far)
# write_csv(matched.so.far, "medicareD_matched_1325.csv")
# write_csv(chem.anti, "medicareD_unmatched_3173.csv")
```

Now have to start altering the Medicare Part D drug names:

```{r}
chem.anti <- chem.anti %>% 
  mutate(alt_name = str_replace(generic_name, "hcl", "hydrochloride"))
hcl.semi <- chem.anti %>% 
  semi_join(atc, by = c("alt_name" = "chem_formula"))
hcl.anti <- chem.anti %>% 
  anti_join(atc, by = c("alt_name" = "chem_formula"))
dim(hcl.semi)
dim(hcl.anti)
matched.so.far <- matched.so.far %>% 
  bind_rows(hcl.semi %>% select(-alt_name))
```

A potential issue is drugs with multiple ingredients. It seems that many drugs have multiple active ingredients separted potentially by "/", ",", or "-" in the `generic_name` column. 

```{r mult ingred}
hcl.anti <- hcl.anti %>% 
  mutate(mult.forsla = str_count(alt_name, "/"),
         mult.comm = str_count(alt_name, ","),
         mult.dash = str_count(alt_name, "-"),
         mult.all = mult.forsla + mult.comm + mult.dash)
nrow(hcl.anti %>% 
       filter(mult.all > 0))
sum(str_detect(hcl.anti$alt_name, "/")) + sum(str_detect(hcl.anti$alt_name, ",")) + sum(str_detect(hcl.anti$alt_name, "-"))
# some names use multiple signs, for some drugs this is a sign of multiple ingredients and for others its just naming convention
```

One "/" notation that stood out for some drug names was "/pf". Upon a little bit of research, it seems that "pf" stands for "pre-filled", as in a pre-filled syringe (such as a vaccine). 

```{r}
slash.pf <- hcl.anti %>% filter(str_detect(alt_name, "pf"))
nrow(slash.pf)
glimpse(slash.pf)
# note that "Pf", "PF", and "pF" bring up no results
nrow(slash.pf %>% filter(str_detect(alt_name, "/pf")))
# all "pf" instances are "/pf" - can remove these and maybe get a few more matches
sum(str_detect(hcl.anti$alt_name, "/"))
hcl.anti <- hcl.anti %>% 
  mutate(alt_name = str_trim(str_replace(alt_name, "/pf", "")),
         mult.forsla = str_count(alt_name, "/"),
         mult.comm = str_count(alt_name, ","),
         mult.dash = str_count(alt_name, "-"),
         mult.all = mult.forsla + mult.comm + mult.dash)
sum(str_detect(hcl.anti$alt_name, "/"))
# some of the drugs still have "/" in the alt_name column
hcl.anti <- hcl.anti %>%
  mutate(alt_name = str_replace(alt_name, "hcl", "hydrochloride"))
# match after pf removal and second "hcl" replacement
pf.semi <- hcl.anti %>% 
  semi_join(atc, by = c("alt_name" = "chem_formula")) # %>% 
  #bind_rows(
   # hcl.anti %>% 
    #  semi_join(atc, by = c("alt_name" = "drugname_generic")) # this result is zero
  #)
dim(pf.semi)
matched.so.far <- matched.so.far %>% 
  bind_rows(pf.semi %>% select(brand_name, generic_name))
dim(matched.so.far)
pf.anti <- hcl.anti %>% 
  anti_join(atc, by = c("alt_name" = "chem_formula"))
# sanity check:
nrow(pf.anti) + nrow(matched.so.far)
```

Okay, now to start tackling some of the multiple-ingredient drugs. The "/" seems to be the easiest to tackle.

```{r fig.height=4, fig.width=4}
table(pf.anti$mult.forsla)
pf.anti %>% 
  ggplot(aes(mult.forsla)) +
    geom_histogram(binwidth = 0.5) +
    xlab("number of slashes")
```

The max number of forward slashes is 7 and 5 or greater is a small enough number of drugs to be looked at:

```{r}
pf.anti %>% filter(mult.forsla >= 5)
```

Many of these seem to be multivitamins and other nutrient supplements. Although care should be taken. A drug called Enlyte, for example, looks on the surface to be a multivitamin but it turns out to be a treatment for certain kinds of depression caused by folate deficiency. 

```{r}
pf.anti.2ingr <- pf.anti %>%
  filter(mult.forsla == 1) %>% 
  select(ends_with("name")) %>% 
  separate(alt_name, c("ingrd_1", "ingrd_2"), sep = "/") %>% 
  gather("ingrd", "chem_name", 3:4) %>%
  mutate(chem_name = str_trim(chem_name)) %>%
  arrange(brand_name)
pf.2ingr.c.semi <- pf.anti.2ingr %>% 
  semi_join(atc, by = c("chem_name" = "chem_formula"))
pf.2ingr.c.anti <- pf.anti.2ingr %>% 
  anti_join(atc, by = c("chem_name" = "chem_formula"))
pf.2ingr.g.semi <- pf.2ingr.c.anti %>% 
  semi_join(atc, by = c("chem_name" = "drugname_generic"))
pf.2ingr.g.anti <- pf.2ingr.c.anti %>% 
  anti_join(atc, by = c("chem_name" = "drugname_generic"))
pf.2ingr.so.far <- pf.2ingr.c.semi %>% 
  bind_rows(pf.2ingr.g.semi)
# matched
dim(pf.2ingr.so.far)
# to go
dim(pf.2ingr.g.anti)
```

Upon further exploration, there seems to be a problem - the `drug_uses.csv` dataset does not seem to contain the full ATC code system. That's a problem.

# Will abandon this attempt to merge the medicare part d data with the `drug_uses.csv` file. Will start over with `atc_codes_clean.csv`.

