---
title: "part_d_atc_codes_clean_csv_merge"
author: "Darya Akimova"
date: "February 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Packages:
```{r packages}
library(tidyverse)
library(data.world)
```

Data:
```{r data}
drug.spend.ds <- "https://data.world/data4democracy/drug-spending"
# drug names from Medicare Part D data:
d.drugs <- data.world::query(
  data.world::qry_sql("SELECT DISTINCT brand_name, generic_name FROM `spending_part_d_2011to2015_tidy`"),
  dataset = drug.spend.ds
  ) %>% 
  mutate_all(str_trim)  # just in case
glimpse(d.drugs)
# atc classification system from atc_codes_clean.csv
atc <- data.world::query(
  data.world::qry_sql("SELECT * FROM atc_codes_clean"),
  dataset = drug.spend.ds
) 
glimpse(atc)
```

A bit of a mess, and too much information for what is needed right now.

```{r}
atc.sml <- atc %>%
  select(level5_code:kegg) %>% 
  arrange(level5)
glimpse(atc.sml)
map_int(map(atc.sml, unique), length)
```

It's still a bit of a mess from where I stopped in my attempt to clean up this data. Wasn't sure how the names would be used.Also note, some compounds have more than 1 level 5 code - which means that they have multiple uses according to the ATC classification.

Getting rid of the KEGG codes in the names would be a good place to start.

```{r}
atc.sml <- atc.sml %>% 
  mutate(lvl5_cln = str_replace(level5, "\\[DG.+\\]", "") %>% str_trim() %>% str_to_lower())
str_sub(atc.sml$kegg, start = 1, end = 6) <- ""
glimpse(atc.sml)
atc.sml <- atc.sml %>% 
  mutate(kegg_cln = str_replace(kegg, "\\([[:upper:]]{2,}.+\\)", "") %>% str_trim() %>% str_to_lower())
atc.sml$kegg_cln[is.na(atc.sml$kegg_cln)] <- atc.sml$lvl5_cln[is.na(atc.sml$kegg_cln)]
atc.cln <- atc.sml %>% 
  select(level5_code, kegg_cln) %>% 
  distinct() %>% 
  arrange(kegg_cln)
dim(atc.sml)
dim(atc.cln)
glimpse(atc.cln)
rm(atc.sml)  # don't need anymore
```

Now that that's been cleaned up a bit, let's try some matching:

```{r}
brand.semi.t1 <- d.drugs %>% 
  semi_join(atc.cln, by = c("brand_name" = "kegg_cln"))
dim(brand.semi.t1)
brand.anti.t1 <- d.drugs %>% 
  anti_join(atc.cln, by = c("brand_name" = "kegg_cln"))
dim(brand.anti.t1)
generic.semi.t1 <- brand.anti.t1 %>% 
  semi_join(atc.cln, by = c("generic_name" = "kegg_cln"))
dim(generic.semi.t1)
generic.anti.t1 <- brand.anti.t1 %>% 
  anti_join(atc.cln, by = c("generic_name" = "kegg_cln")) %>% 
  arrange(generic_name)
dim(generic.anti.t1)
matched.so.far <- brand.semi.t1 %>% 
  mutate(key = brand_name) %>% 
  bind_rows(
    generic.semi.t1 %>% 
      mutate(key = generic_name)
    )
dim(matched.so.far)
# break 1
# write_csv(matched.so.far, "partD_matched_2034_pass1.csv")
# write_csv(generic.anti.t1, "partD_unmatched_2464_pass1.csv")
```

This is much better than the first pass with `drug_uses.csv`, where I only got 1325 matches on the first pass. A good sign! 

Now, drugs with multiple ingredients:

```{r}
generic.anti.t1 <- generic.anti.t1 %>% 
  mutate(
    key = generic_name,
    mult.slash = str_count(key, "/"),
    mult.comm = str_count(key, ","),
    mult.dash = str_count(key, "-"),
    mult.all = mult.slash + mult.comm + mult.dash
    )
glimpse(generic.anti.t1)
nrow(generic.anti.t1 %>% filter(mult.all > 0))
# from previous attempts at matching: "/pf" probably indicates a prefilled syringe
slash.pf <- generic.anti.t1 %>% filter(str_detect(key, "pf"))
nrow(slash.pf)
glimpse(slash.pf)
pf.semi <- generic.anti.t1 %>% 
  mutate(key = str_trim(str_replace(key, "/pf", ""))) %>% 
  semi_join(atc.cln, by = c("key" = "kegg_cln")) %>% 
  select(brand_name:key)
dim(pf.semi)
matched.so.far <- matched.so.far %>% 
  bind_rows(pf.semi)
pf.anti <- generic.anti.t1 %>% 
  mutate(key = str_trim(str_replace(key, "/pf", ""))) %>% 
  anti_join(atc.cln, by = c("key" = "kegg_cln"))
dim(pf.anti)
# another from previous attempts:
sum(str_detect(pf.anti$key, "hcl"))
sum(str_detect(atc.cln$kegg_cln, "hcl"))
sum(str_detect(atc.cln$kegg_cln, "hydrochloride"))
# "hcl" = "hydrochloride"
pf.anti <- pf.anti %>% 
  mutate(key = str_replace_all(key, "hcl", "hydrochloride"),
         # same principle as hcl, "hbr" = "hydrobromide"
         key = str_replace_all(key, "hbr", "hydrobromide"))
hcl.semi <- pf.anti %>% 
  semi_join(atc.cln, by = c("key" = "kegg_cln"))
dim(hcl.semi)
glimpse(hcl.semi %>% select(brand_name:key))
# 431 more matches!
matched.so.far <- matched.so.far %>% 
  bind_rows(hcl.semi %>% select(brand_name:key))
dim(matched.so.far)
dim(unique(matched.so.far))
# nice, no accidental repeats
hcl.anti <- pf.anti %>% 
  anti_join(atc.cln, by = c("key" = "kegg_cln"))
dim(hcl.anti)
hcl.anti <- hcl.anti %>% 
  mutate(mult.slash = str_count(key, "/"),
         mult.all = mult.slash + mult.comm + mult.dash)
nrow(hcl.anti %>% filter(mult.all > 0))
# a little bit of cleanup
rm(brand.semi.t1, brand.anti.t1, generic.anti.t1, generic.semi.t1, pf.semi, pf.anti, slash.pf, hcl.semi)
```

Up to 1735 of the remaining drugs may have multiple ingredients. However, not necessarily all of them are because the characters `,` and `-` especially have different uses in drug names. The `/` character is the most straightforward currently to tackle. 

```{r}
table(hcl.anti$mult.slash)
all.slash <- hcl.anti %>% 
  filter(mult.slash > 0)
no.slash <- hcl.anti %>% 
  filter(mult.slash == 0)
dim(all.slash)
dim(no.slash)
# for drugs/items not in the ATC system
matched.other <- no.slash %>% 
  filter(
    key == "alcohol antiseptic pads" | key == "gauze bandage" | str_detect(key, "needle") | str_detect(key, "syring") | str_detect(key, "syringe") | str_detect(key, "insulin admin. supplies") | str_detect(key, "insulin pump cartridge") | str_detect(key, "lancets")
    ) %>% 
  select(brand_name:key)
other.anti <- no.slash %>% 
  anti_join(matched.other, by = "key")
dim(matched.so.far)
dim(matched.other)
dim(other.anti)
dim(all.slash)
# still adds up to 4498, good
other.anti <- other.anti %>% 
  mutate(key = str_replace_all(key, "levonorgestrel-ethin estradiol", "levonorgestrel and ethinylestradiol"),
         key = str_replace_all(key, "acetaminophen with codeine", "codeine and paracetamol"),
         key = str_replace_all(key, "colistin (colistimethate na)", "colistin"),
         key = str_replace_all(key, "cyanocobalamin (vitamin b-12)", "cyanocobalamin"),
         key = str_replace_all(key, "desogestrel-ethinyl estradiol", "desogestrel and ethinylestradiol"),
         key = str_replace_all(key, "doxycycline monohydrate", "doxycycline"),
         key = str_replace_all(key, "ergocalciferol (vitamin d2)", "ergocalciferol"),
         key = str_replace_all(key, "erythromycin base", "erythromycin"),
         key = str_replace_all(key, "ethynodiol d-ethinyl estradiol", "etynodiol and ethinylestradiol"),
         key = str_replace_all(key, "norethindrone-e.estradiol-iron", "norethisterone and ethinylestradiol/iron"),
         key = str_replace_all(key, "norethindrone-ethinyl estrad", "norethisterone and ethinylestradiol"),
         key = str_replace_all(key, "norethindrone ac-eth estradiol", "norethisterone and ethinylestradiol"),
         key = str_replace_all(key, "norethindrone-mestranol", "norethisterone and estrogen"),
         key = str_replace_all(key, "norgestimate-ethinyl estradiol", "norgestimate and ethinylestradiol"),
         key = str_replace_all(key, "norgestrel-ethinyl estradiol", "norgestrel and ethinylestradiol"),
         key = str_replace_all(key, "ondansetron hydrochloride", "ondansetron")
         ) %>% 
  mutate_all(str_trim)
other.rpl.semi <- other.anti %>% 
  semi_join(atc.cln, by = c("key" = "kegg_cln"))
other.rpl.anti <- other.anti %>% 
  anti_join(atc.cln, by = c("key" = "kegg_cln"))
dim(other.rpl.semi)
dim(other.rpl.anti)
matched.so.far <- matched.so.far %>% 
  bind_rows(other.rpl.semi %>% select(brand_name:key))
rm(other.anti, other.rpl.semi)
other.rpl.anti <- other.rpl.anti %>% 
  mutate(mult.slash = str_count(key, "/"),
         mult.comm = str_count(key, ","),
         mult.dash = str_count(key, "-"),
         mult.all = mult.slash + mult.comm + mult.dash)
all.slash <- all.slash %>% 
  bind_rows(other.rpl.anti %>% filter(mult.slash > 0))
other.rpl.anti <- other.rpl.anti %>% 
  filter(mult.slash == 0)
```

The major variables to keep track of now:
- `matched.so.far` - list of successfully matched drugs with a `key` column that they can be matched by to the atc dataset
- `matched.other` - "matched" medical items not in atc dataset, such as needles, syringes, and ethanol pads
- `all.slash` - drugs definitely containing multiple ingredients separated by `/`
- `other.rpl.anti` - misc drugs that need to be matched, may not be in atc dataset

```{r}
# sanity checks
nrow(matched.so.far) + nrow(matched.other) + nrow(all.slash) + nrow(other.rpl.anti)
nrow(unique(matched.so.far)) + nrow(unique(matched.other)) + nrow(unique(all.slash)) + nrow(unique(other.rpl.anti))
# break 2
# write_csv(bind_rows(matched.so.far, matched.other), "partD_matched_2773_break2.csv")
# all.slash %>% 
#   select(brand_name:key) %>% 
#   bind_rows(
#     other.rpl.anti %>% 
#       select(brand_name:key)
#     ) %>% 
# write_csv("partD_unmatched_1725_break2.csv")

table(all.slash$mult.slash)
mult.ingred.5 <- all.slash %>% 
  filter(mult.slash >= 4) %>% 
  arrange(generic_name)
glimpse(mult.ingred.5)
# some are vitamin complexes, but some are drugs - will think about what to do with these
# start with 2 ingredient drugs, separated by only 1 `/`
gen.anti.2ingrd <- all.slash %>% 
  filter(mult.slash == 1) %>% 
  mutate(alt_copy = key) %>% 
  separate(alt_copy, c("ingrd_1", "ingrd_2"), sep = "/") %>% 
  select(brand_name:key, starts_with("ingrd")) %>% 
  mutate_all(str_trim)
glimpse(gen.anti.2ingrd)
gen.anti.2ingrd <- gen.anti.2ingrd %>% 
  gather("ingrd", "chem_name", 4:5) %>% 
  arrange(key)
glimpse(gen.anti.2ingrd)
ingrd2.semi.t1 <- gen.anti.2ingrd %>% 
  semi_join(atc.cln, by = c("chem_name" = "kegg_cln"))
dim(ingrd2.semi.t1)
ingrd2.semi.matched <- ingrd2.semi.t1 %>% 
  group_by(brand_name, generic_name) %>% 
  count() 
table(ingrd2.semi.matched$n)
# the n = 2 drugs represent those that have been fully matched, the drugs with a n = 1 have one ingredient unmatched
ingrd2.anti.t1 <- gen.anti.2ingrd %>% 
  anti_join(atc.cln, by = c("chem_name" = "kegg_cln")) %>% 
  arrange(chem_name)
dim(ingrd2.anti.t1)
dim(ingrd2.anti.t1 %>% select(chem_name) %>% unique())
test <- ingrd2.anti.t1 %>% select(chem_name) %>% unique() %>% arrange(chem_name)
```

